Long-Term Reproducibility through Replication:\\ the ReScience Initiative

### 2017-06-30 draft.tex
fixed typos

Confronted with this problem, a small but growing number of journals and
publishers have reacted by adopting explicit policies for data and software, for example the PLOS instructions on
\href{http://journals.plos.org/plosone/s/materials-and-software-sharing}{Materials
  and Software Sharing} and
on \href{http://journals.plos.org/plosone/s/data-availability}{Data Availability},
and the
\href{https://elifesciences.org/elife-news/inside-elife-forking-software-used-elife-papers-github}{recent
  annoucement} by eLife on forking (creating a linked copy of) software used in eLife papers to GitHub.
Such policies help to ensure access to code and data in a well-defined format
\citep{perkel:2016} but this will not guarantee reproducibility nor
correctness.
At the educational and methodological levels, things have started to change with a
growing literature on best practices for making computations reproducible
\citep{sandve:2013, crook:2013, wilson:2014, halchenko:2015, janz:2015,
  hinsen:2015}.
  Related initiatives such as Software and Data Carpentry
\citep{wilson:2016} are of note since their goal is {\em to make
scientists more productive, and their work more reliable, by teaching them basic computing skills}.
Such best practices could be applied to already published research codebases as well, provided the original authors are willing to take on the challenge of 
re-implementing their software for the sake of better science.
Unfortunately, this is unlikely since the
incentives for doing such time-consuming work are low or nonexistent.
Furthermore, if the original authors made mistakes in their original
implementation, chances are that they will reproduce their mistakes.\\
% (just an educated guess, we have no data supporting this assertion)
%% At this point, the question is what do we do? Do we have to declare
%% the research to be lost once and for all because the associated software or
%% code is nowhere to be found / lost / damaged / malfunctioning / not runnable /
%% not usable / does not give original results / etc. We would doom ourselves by
%% doing so

@@

Today's electronic publishing platforms for scientific research still
show their origins in paper-based publishing. Except for the
replacement of printed paper by a printable PDF file, not much has
changed. Although it is increasingly realized that software and data
should be integral parts of most scientific publications today, they
are at best relegated to the status of ``supplementary material'', and
systematically excluded from the peer review process. In fact, to the
best of our knowledge, ReScience is the only scientific journal that
aims to verify the correctness of scientific software. As our
experience has shown, it is far easier to graft publication onto a
software development platform than to integrate software reviewing
into a publishing platform.

>>>>

Confronted with this problem, a small but growing number of journals and
publishers have reacted by adopting explicit policies for data and software, for example the PLOS instructions on
\href{http://journals.plos.org/plosone/s/materials-and-software-sharing}{Materials
  and Software Sharing} and
on \href{http://journals.plos.org/plosone/s/data-availability}{Data Availability},
and the
\href{https://elifesciences.org/elife-news/inside-elife-forking-software-used-elife-papers-github}{recent
  annoucement} by eLife on forking (creating a linked copy of) software used in eLife papers to GitHub.
Such policies help to ensure access to code and data in a well-defined format
\citep{perkel:2016} but this will not guarantee reproducilbility nor
correctness.
At the educational and methodological levels, things have started to change with a
growing literature on best practices for making computations reproducible
\citep{sandve:2013, crook:2013, wilson:2014, halchenko:2015, janz:2015,
  hinsen:2015}.
  Related initiatives such as Software and Data Carpentry
\citep{wilson:2016} are of note since their goal is {\em to make
scientists more productive, and their work more reliable, by teaching them basic computing skills}.
Such best practices could be applied to already published research codebases as well, provided the original authors are willing to take on the challenge of 
re-implementing their software for the sake of better science.
Unfortunately, this is unlikely since the
incentives for doing such time-consuming work are low or nonexistent.
Furthermore, if the original authors made mistakes in their original
implementation, chances are that they will reproduce their mistakes.\\
% (just an educated guess, we have no data supporting this assertion)
%% At this point, the question is what do we do? Do we have to declare
%% the research to be lost once and for all because the associated software or
%% code is nowhere to be found / lost / damaged / malfunctioning / not runnable /
%% not usable / does not give original results / etc. We would doom ourselves by
%% doing so

@@

Today's electronic publishing platforms for scientific research still
show their origins in paper-based publishing. Except for the
replacement of printed paper by a printable PDF file, not much has
changed. Although it is increasingly realized that software and data
should be integral parts of most scientific publications today, they
are at best relegated to the status of ``supplementary material'', and
systematically excluded from the peer review process. In fact, to the
best of our knowledge, ReScience is the only scientific journal that
aims to verify the correctness of scientific software. As our
experience has shown, it is far easier to graft publication onto a
software development platform than to integrate software reviewing
into a publishing platform. Furthermore, tools that will allow for the automated
validation of computational models and the automated verification of correctness are being actively 
developed in the community (see e.g. 
\href{https://github.com/scidash/sciunit}{SciUnit} or 
\href{https://github.com/OpenSourceBrain/osb-model-validation}{OSB-model-validation}). 
An integration of such frameworks, which would greatly enhance the 
verification and validation process, seems feasible for the existing software 
development platform.
